---
title: "Mortality and survival in Game of Thrones"
teaching: 15
exercises: 40
questions:
- "How can I perform preliminary data quality visualisations?"
objectives:
- Learn how to undertake simple dataset review and visualitions
- Perform suvival analyses
keypoints:
- "Load data into R."
- "Perform basic data visualisations using ggplot2"
source: Rmd
---

```{r, include=FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("17-")
```

Let's start by downnloading Game of Thrones characters' mortality data from [here](https://figshare.com/articles/Game_of_Thrones_mortality_and_survival_dataset/8259680?mc_cid=6ee60dc1ef&mc_eid=f10fe3b3f2).

There is an additional metadata encoding table that we have prepared, which you can download from [here]()

> ## Challenge 1
>
> Save all three files your `data/` directory and change the working directory to it.
> Now read the `data/character_data_S01-S08.csv` file into R.
>
> > ## Solution to Challenge 1
> >
> > ```{r, eval=FALSE}
> > setwd('path/to/data')
> > got_dat <- read.csv(file = "character_data_S01-S08.csv", header = TRUE, stringsAsFactors = FALSE)
> > meta <- read.csv(file = "encoding.csv", header = TRUE, stringsAsFactors = FALSE)
> > ```
> {: .solution}
{: .challenge}

```{r, eval=TRUE, echo=FALSE}
got_dat <- read.csv(file = "./data/character_data_S01-S08.csv", header = TRUE, stringsAsFactors = FALSE)
meta <- read.csv(file = "./data/encoding.csv", header = TRUE, stringsAsFactors = FALSE)
```

Once that data is loaded, let's evualuate its quality.

> ## Challenge 2
>
> Does the table with GoT characters's mortality data look reasonable?
>
> > ## Solution to Challenge 2
> >
> > ```{r, eval=TRUE}
> > ## make a summary for each column 
> > summary(got_dat)
> > ## check for NAs in each column
> > apply(got_dat, 2, function(x) all(is.na(x)))
> > ```
> {: .solution}
{: .challenge}
